---
title: "Assignment 1"
author: "Elliot Gavrin"
date: "2023-01-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
`r knitr::spin_child("dataGenerato.R")`
dataGenerato.R generates random distributed data with no negative or 0 values to be used as values for limb length and limb width.

`r knitr::spin_child("volumeEstimato.R")`
volumeEstimato.R calculates limb volume from the generated limb length and limb width

1.
```{r}
MyData <-read.csv("C:/Users/egavr/OneDrive/Documents/BIOL432/a1/measurements.csv")
dim(MyData)
head(MyData)
tail(MyData)
str(MyData)
summary(MyData)
library(data.table)
library(dplyr)
library(ggplot2)
```

2.
```{r}
DataSort <- data.frame(MyData$Organisms, MyData$Observer, MyData$Limb_volume) #creates data frame
```

3. 

```{r}
Species1 <- subset(DataSort, Organisms == "Poecile atricapillus")
Species1Avg <- mean(Species1$MyData.Limb_volume)

Species2 <- subset(DataSort, Organisms == "Columba livia")
Species2Avg <- mean(Species2$MyData.Limb_volume)

Species3 <- subset(DataSort, Organisms == "Cyanocitta cristata")
Species3Avg <- mean(Species3$MyData.Limb_volume)

Species4<- subset(DataSort, Organisms == "Pica serica")
Species4Avg <- mean(Species4$MyData.Limb_volume)

Species5 <- subset(DataSort, Organisms == "Anas acuta")
Species5Avg <- mean(Species5$MyData.Limb_volume)

dfavg <-data.frame(Species1Avg, Species2Avg, Species3Avg, Species4Avg, Species5Avg)
colnames(dfavg) = c('Poecile atricapillus','Columba livia','Cyanocitta cristata','Pica serica','Anas acuta')
rownames(dfavg) = c('Average Limb Volume')
averagevolume <-table(dfavg)
print(dfavg)
```

4.
```{r}
observnum <-rename(count(DataSort, MyData.Organisms, MyData.Observer), Freq = n)
setDT(observnum)
print(observnum)
```

5.
```{r}
DataSort <- data.frame(MyData$Organisms, MyData$Observer, MyData$Limb_volume)
ggplot(aes(y=MyData.Limb_volume,color=MyData.Organisms), data=DataSort)+
  geom_boxplot(bins=20)+
  labs(x = "Species", y = "Limb volume", color = "Species")+
  theme_classic()
```

Figure 1: Box plot displaying the distribution of limb volume in each of the five different species: Anas acuta, Columba livia, Cyanocitta cristata, Pica serica, and Poecile atripillus.

6.
```{r}
ggplot(aes(x=Freq, color=MyData.Observer), data=observnum)+
  geom_bar(bins=10)+
  labs(x = "Frequency", y="", color="Observer")+
  facet_wrap(vars(MyData.Organisms), scales='free')+
  theme_classic()
```

Figure 2: Multi-panel plot showing frequency histograms for each of the 5 species. Colored outlines around the bins represent the individual observer's count. 
